### 2020-12-02 latex/acrobot.tex
Finished first version of rotation and omega proof

\begin{thm}\label{thm:acrobot-energy-stabilization}
    Consider the simple acrobot ~\eqref{eqn:acrobot-hamiltonian} constrained by
    the VNHC ~\eqref{eqn:acrobot-constraint}, whose
    constraint manifold is \(\Gamma \simeq \SxR\).
    Let 
    \[
        E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
        ,
    \]
    be the energy of the simple pendulum obtained by setting \(I = 0\).
\begin{enumerate}
    \item There exists \(I > 0\) small enough that 
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \[
        \mathcal{O} := \left\{(q_u,p_u) \in \Gamma 
        \mid E(q_u,p_u) < E(\pi,0) \right\}
        .
    \]
    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}\).
\item Define \(b : \SxR_{> 0} \rightarrow \R\) by
    \[
        b(\beta,\rho_0) := 
        \frac{5m^2 g l^3 \bar{q}_a \left(
            m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
            - c_\beta\rho_0^2
        \right)}{
        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
        }
        ,
    \]
    and fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
    Suppose \(\int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma > 0\) for all 
    \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
    Then there exists \(I > 0\) small enough that
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \[
        \Omega := \left\{(q_u,p_u) \in \Gamma 
        \mid E(q_u,p_u) < E(0,\bar{\rho})\right\}
        .
    \]
    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega\).
\end{enumerate}
\end{thm}

@@

Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
being \(O(I^2)\).
Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(R(\pi,\underbar{\mu},I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
\[
    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
    .
\]
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for\(\mu_0 \in [\epsilon,\pi[\) we have
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
Choosing \(I^\star \in\, ]0,\min\left\{I_2,I_3\right\}]\) so that
\(I^\star K Q(\epsilon) - (I^\star)^2 r > 0\)
means that
\( P_\mathcal{O}(\mu_0) > \mu_0\)
for all \(\mu_0 \in [\epsilon,\pi[\).

@@

The function \(b(\beta,\rho_0)\) is even and
\(2\pi\)-periodic in \(\beta\), which implies that
\begin{align*}
    \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0) &=
    \int \limits_0^{\sign{\rho_0}2\pi} b(\sigma,\rho_0)d\sigma
    , \\
     &=\sign{\rho_0}\int\limits_0^{2\pi}b(\sigma,\rho_0)d\sigma
     , \\
     &= \sign{\rho_0} \hat{\rho}_1(2\pi,\rho_0)
     .
\end{align*}
Thus, we have that
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I\sign{\rho_0}\hat{\rho}_1(2\pi,\rho_0) 
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

Note that \(b(\beta,\rho_0)\) is even in \(\rho_0\), which means
\(\hat{\rho}_1(2\pi,\rho_0) = \hat{\rho}_0(2\pi,\abs{\rho_0})\).
Recalling the assumption that there exists \(\bar{\rho} > 0\) so that
\(\int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma > 0\) for all 
\(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\),
we have that \(\hat{\rho}_1(2\pi,\rho_0) > 0\)
for every \(\rho_0\) with \(\sqrt{60m^2gl^3} < \abs{\rho} \leq \bar{rho}\).
Finally, \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\). 
Defining 
\[
    D := [-\bar{\rho},\sqrt{60m^2gl^3}[ \, \cup \, ]\sqrt{60m^2gl^3},\bar{\rho}]
    ,
\]
these facts all together imply that, for
each \(\rho_0 \in D\),
there exists \(I > 0\) small enough that 
\[
    \begin{cases}
        P_\mathcal{R}(\rho_0) > \rho_0 & \rho_0 > 0
        ,\\
        P_\mathcal{R}(\rho_0) < \rho_0 & \rho_0 < 0
        .\\
    \end{cases}
\]

\subsubsection*{Energy Gain on \(\mathcal{R}\)}

The Poincar\'{e} section \(P_\mathcal{R}\) allows us understand the evolution of
\(\hat{\rho}(\beta,\rho_0,I)\) by studying the evolution of the discrete time
system
\begin{equation}\label{eqn:rhohat-discrete}
    \begin{cases}
        \rho(n+1) := P_\mathcal{R}(\rho(n)) 
        = \rho(n) + I\sign{\rho_0}\hat{\rho}_1(2\pi,\rho(n)) + 
            R(\sign{\rho_0}2\pi,\rho(n),I)
        , \\
        \rho(0) = \rho_0
        .
    \end{cases}
\end{equation}
Tolerating the abuse of notation, \(\rho(n)\) represents the distance along the
\(p_u\)-axis when an orbit of the constrained dynamics intersects the
\(p_u\)-axis for the \(n\)th time, assuming the orbit was initialized at
\((q_u,p_u) = (0,\rho_0)\).

We wish to show \(\abs{\rho(n)}\) eventually reaches \(\bar{\rho}\).
That is, we wish to find some \(I^\star \in \,]0,I_1]\) making
\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all 
\(\rho_0 \in D\)
We will restrict ourselves, without loss of generality, to the case where
\(\rho_0 > 0\).

Recall that \(R(\beta,\rho_0,I)\) is \(O(I^2)\) and smooth in all
its parameters.
Therefore, it is bounded below on the compact set
\(\rho_0 \in [\sqrt{60m^2gl^3},\bar{\rho}]\) by some value
\(R(2\pi,\underbar{\rho},I)\).
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_2 > 0\) and \(r > 0\) so that, for all \(I \in [-I_2,I_2]\),
\[
    R(2\pi,\rho_0,I) \geq R(2\pi,\underbar{\rho},I) > -I^2r
    .
\]

%TODO: replace this with an analysis of poincare section
Let now \(\bar{\rho} > \sqrt{60m^2gl^3}\) and take the compact set
\[
    \bar{\mathcal{R}} := \left\{(q_u,p_u) \in \Gamma 
    \mid E(\pi,0) \leq E(q_u,p_u) \leq E(0,\bar{\rho}) \right\}
    .
\]
The (modified) remainder term \(\sign{\rho_0}{R}(\beta,\rho_0,I)\) 
is smooth and \(O(I^2)\), which means
it attains its minimum \(I^2\hat{r}\) somewhere \(\bar{\mathcal{R}}\).
Likewise, \(\hat{\rho}_1(2\pi,\abs{\rho_0})\) also attains its minimum, which we
denote by \(\underbar{\rho}_1\).
We cannot adimensionalize \(b(\beta,\rho_0)\), so refer to one of the 
assumptions of Theorem \ref{thm:acrobot-energy-stabilization}
to suppose  that \(\hat{\rho}_1(2\pi,\rho_0) > 0 \) for all 
\(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\).
In particular, this means \(\underbar{\rho}_1 > 0\).

There exists \(I > 0\) making 
\(I\underbar{\rho}_1 + I^2\hat{r} > 0\).
Hence,
\begin{align*}
    \abs{\rho_0} + I\hat{\rho}_1(2\pi,\abs{\rho_0}) + \sign{\rho_0}R(2\pi,\rho_0,I)
    &\geq \sqrt{60m^2gl^3} + I\underbar{\rho}_1 + I^2\hat{r}
    \\
    &> \sqrt{60m^2gl^3}
    \\
    &> 0
    ,
\end{align*}
which implies that
\begin{align*}
    \abs{P_\mathcal{R}(\rho_0)} - \abs{\rho_0}
    &= \abs{ \abs{\rho_0} + I\hat{\rho}_1(2\pi,\rho_0) +
    \sign{\rho_0}R(2\pi,\rho_0,I)} - \abs{\rho_0}
    \\
    &= \abs{\rho_0} + I\hat{\rho}_1(2\pi,\rho_0) +
    \sign{\rho_0}R(2\pi,\rho_0,I) - \abs{\rho_0}
    \\
    &= I\hat{\rho}_1(2\pi,\rho_0) + \sign{\rho_0}R(2\pi,\rho_0,I)
    \\
    &> 0
    .
\end{align*}

We conclude that there exists \(I^\prime(\bar{\rho})\)
small enough that, for all \( 0 < I \leq I^\prime(\bar{\rho})\), 
the Poincar\'{e} section \(P_\mathcal{R}(\rho_0)\) is monotonically increasing
when \(\sqrt{60m^2gl^3} < \rho_0 < \bar{\rho}\) and
monotonically decreasing when \(-\bar{\rho} < \rho_0 < -\sqrt{60m^2gl^3}\).

\subsection*{Energy Gain on \(\Omega\)}
We will prove the acrobot gains energy on the set 
\[
    \Omega := \left\{ (q_u,p_u) \in \Gamma 
     \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    .
\]
% Can we use the jacobian? The spectrum is hard to find.
% TODO: W and D's boundary is two elements: the point (pi,0) and the curve of the
% homoclinic orbit connecting (pi,0) to (-pi,0). We need to show that almost
% every initial condition starting in \(mathcal{O}\) will NOT converge to (pi,0), and
% therefore exit D and enter W.

Define the set of points in \(\mathcal{R}\) which have pseudo-radius in \(U\) by
\[
    W := \left\{(q_u,p_u) \in \Gamma \mid |\rho(q_u,p_u)| \in U \right\}
    ,
\]
as in Figure \ref{fig:acrobot-rot-W}.
Monotonicity of \(P(\rho_0)\) means there are no closed orbits of 
\((q_u,p_u)\) on \(W\). 
Furthermore, \(|p_u|\) is increasing every time the orbit hits the
\(p_u\)-axis.

\begin{figure}
    \centering
    % TODO: Display the set W, which is the two rotation bands at the top and
    % bottom with outer limit +- \bar{mu}  and inner limit +- sqrt{60m^2gl^3}.
    % Show also that the inner limit connects to D
    %\includestandalone[width=0.5\textwidth]{images/acrobot_rot_W}
    \caption{The set \(W\) of points which have pseudo-radius 
        \(|\rho| \in ]\sqrt{60m^2gl^3}, \bar{\rho}[\). The inner boundary of \(W\)
        is \((q_u,p_u) = (\pi,0)\), which is also the boundary of the
        oscillation domain \(mathcal{O}\).}
    \label{fig:acrobot-rot-W}
\end{figure}

We conclude that \(|\rho(q_u,p_u)|\) will eventually reach \(\bar{\rho}\). 
We cannot claim this fits the definition  of energy injection 
(since \((0,0) \notin W\)), but we can still say that \((q_u,p_u)\) will
eventually depart \(W\) through the outer boundary in finite time.
The acrobot is thereby gaining energy in the literal sense that its momentum
is increasing on \(W\).

If we set \(I < 0\), we have \(P(|\rho_0|) < |\rho_0|\).
Hence, \(|p_u|\) is decreasing every time the orbit hits the \(p_u\)-axis, so
the acrobot will eventually depart \(W\) through the inner boundary. 
The acrobot is losing momentum, and therefore it is losing energy on \(W\).

A hybrid supervisor that flips the sign of \(I\) based on the current
value of \(\mu\) will stabilize any energy level of oscillation.
As was the case in the VLP, this supervisor should implement some kind of
hysteresis to avoid infinite switching.

>>>>

\begin{thm}\label{thm:acrobot-energy-stabilization}
    Consider the simple acrobot ~\eqref{eqn:acrobot-hamiltonian} constrained by
    the VNHC ~\eqref{eqn:acrobot-constraint}, whose
    constraint manifold is \(\Gamma \simeq \SxR\).
    Let 
    \[
        E(q_u,p_u) := \frac{p_u^2}{10ml^2} + 3mgl(1 - \cos(q_u))
        ,
    \]
    be the energy of the simple pendulum obtained by setting \(I = 0\).
\begin{enumerate}
    \item There exists \(I > 0\) small enough that 
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \[
        \mathcal{O} := \left\{(q_u,p_u) \in \Gamma 
        \mid E(q_u,p_u) < E(\pi,0) \right\}
        .
    \]
    If instead \(I < 0\), the VNHC dissipates energy on \(\mathcal{O}\).
\item Define \(b : \SxR_{> 0} \rightarrow \R\) by
    \[
        b(\beta,\rho_0) := 
        \frac{5m^2 g l^3 \bar{q}_a \left(
            m^2gl^3\left(18s_\beta^2 + 30c_\beta(1 - c_\beta)\right)
            - c_\beta\rho_0^2
        \right)}{
        |\rho_0|\sqrt{\rho_0^2 - 30m^2gl^3(1 - c_\beta)}
        }
        ,
    \]
    and set \(S(\rho_0) := \int \limits_{0}^{2\pi} b(\sigma,\rho_0)d\sigma\).
    Fix \(\bar{\rho} > \sqrt{60m^2gl^3}\).
    Suppose there exists \(\epsilon > 0\) so that \(S(\rho_0) > \epsilon\) for
    all \(\rho_0 \in \, \left]\sqrt{60m^2gl^3}, \bar{\rho}\right]\).
    Then there exists \(I > 0\) small enough that
    ~\eqref{eqn:acrobot-constraint} injects energy into the acrobot on
    \[
        \Omega := \left\{(q_u,p_u) \in \Gamma 
        \mid E(q_u,p_u) < E(0,\bar{\rho})\right\}
        .
    \]
    If instead \(I < 0\), the VNHC dissipates energy on \(\Omega\).
\end{enumerate}
\end{thm}

@@

Next, recall that \(R(\pi,\mu_0,I)\) is smooth in all its parameters as well as
being \(O(I^2)\).
Therefore, it is bounded below on the compact set \(\mu_0 \in [\epsilon,\pi]\)
by some value \(R(\pi,\underbar{\mu},I)\). 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some
\(I_3 > 0\) and \(r > 0\) so that, for all \(I \in [-I_3, I_3]\),
\[
    R(\pi,\mu_0,I) \geq R(\pi, \underbar{\mu}, I) > -I^2 r
    .
\]
Furthermore, \(Q(\mu_0)\) is a strictly increasing function, so
for\(\mu_0 \in [\epsilon,\pi[\) we have
\[
    P_\mathcal{O}(\mu_0) > \mu_0 + I K Q(\epsilon) - I^2 r
    .
\]
Choosing \(I^\star \in\, ]0,\min\left\{I_2,I_3\right\}]\) so that
\(I^\star K Q(\epsilon) - (I^\star)^2 r \geq 0\)
means that
\( P_\mathcal{O}(\mu_0) > \mu_0\)
for all \(\mu_0 \in [\epsilon,\pi[\).

@@

The function \(b(\beta,\rho_0)\) is even and
\(2\pi\)-periodic in \(\beta\), which implies that
\begin{align*}
    \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0) &=
    \int \limits_0^{\sign{\rho_0}2\pi} b(\sigma,\rho_0)d\sigma
    , \\
     &=\sign{\rho_0}\int\limits_0^{2\pi}b(\sigma,\rho_0)d\sigma
     , \\
     &= \sign{\rho_0} \hat{\rho}_1(2\pi,\rho_0)
     .
\end{align*}
Defining \(S(\rho_0) := \hat{\rho}_1(2\pi,\rho_0)\),
we get
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I\sign{\rho_0}S(\rho_0) 
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

Note that \(b(\beta,\rho_0)\) is even in \(\rho_0\), which means
\(S(\rho_0) = S(\abs{\rho_0})\).
Recall the assumption that there exist \(\bar{\rho} > 0\) and
\(\epsilon > 0\) where \(S(\rho_0) \geq \epsilon\) for all 
\(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\).
Hence, we have that \(S(\rho_0) \geq \epsilon\)
for every \(\rho_0\) with \(\sqrt{60m^2gl^3} < \abs{\rho} \leq \bar{\rho}\).
Finally, \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\). 
Defining 
\[
    D := [-\bar{\rho},\sqrt{60m^2gl^3}[ \, \cup \, ]\sqrt{60m^2gl^3},\bar{\rho}]
    ,
\]
these facts all together imply that, for
each \(\rho_0 \in D\),
there exists \(I > 0\) small enough that 
\[
    \begin{cases}
        P_\mathcal{R}(\rho_0) > \rho_0 & \rho_0 > 0
        ,\\
        P_\mathcal{R}(\rho_0) < \rho_0 & \rho_0 < 0
        .\\
    \end{cases}
\]
As with oscillations, this \(I\) may not work for all \(\rho_0 \in D\). We deal
with this issue next.

\subsubsection*{Energy Gain on a Subset of \(\mathcal{R}\)}

The Poincar\'{e} section \(P_\mathcal{R}\) allows us understand the evolution of
\(\hat{\rho}(\beta,\rho_0,I)\) by studying the evolution of the discrete time
system
\begin{equation}\label{eqn:rhohat-discrete}
    \begin{cases}
        \rho(n+1) := P_\mathcal{R}(\rho(n)) 
        = \rho(n) + I\sign{\rho_0}S(\rho(n)) + 
            R(\sign{\rho_0}2\pi,\rho(n),I)
        , \\
        \rho(0) = \rho_0
        .
    \end{cases}
\end{equation}
Tolerating the abuse of notation, \(\rho(n)\) represents the distance along the
\(p_u\)-axis when an orbit of the constrained dynamics intersects the
\(p_u\)-axis for the \(n\)th time, assuming the orbit was initialized at
\((q_u,p_u) = (0,\rho_0)\).

We wish to show \(\abs{\rho(n)}\) eventually reaches \(\bar{\rho}\).
That is, we wish to find some \(I^\star \in \,]0,I_1]\) making
\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all 
\(\rho_0 \in D\)

We begin with examining the case where \(\rho_0 > 0\).
Recall that \(R(\beta,\rho_0,I)\) is \(O(I^2)\) and smooth in all
its parameters.
Therefore, it is bounded below on the compact set
\(\rho_0 \in [\sqrt{60m^2gl^3},\bar{\rho}]\) by some value
\(R(2\pi,\underbar{\rho},I)\).
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_2 > 0\) and \(r > 0\) so that, for all \(I \in [-I_2,I_2]\),
\[
    R(2\pi,\rho_0,I) \geq R(2\pi,\underbar{\rho},I) > -I^2r
    .
\]
Since \(S(\rho_0) \geq \epsilon\), we find that
\[
    P_\mathcal{R}(\rho_0) > \rho_0 + I\epsilon -I^2r
    ,
\]
for all \(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\).
Choosing \(I_3 \in ]0,I_2]\) so that 
\(I_3\epsilon - (I_3)^2 r \geq 0\) means that
\(P_\mathcal{R}(\rho_0) > \rho_0\)
for all \(\rho_0 \in ]\sqrt{60m^2gl^3},\bar{\rho}]\)

We now examine the case when \(\rho_0 < 0\), where we wish to show that 
\(P_\mathcal{R}(\rho_0) < \rho_0\).
The remainder \(R(\beta,\rho_0,I)\) is bounded above on the compact set 
\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}]\), and 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_4 > 0\) and \(r^\prime > 0\) so that, for all \(I \in [-I_4,I_4]\), 
\[
    R(-2\pi,\rho_0,I) < I^2 r^\prime
    .
\]
Hence,
\begin{align*}
    P_\mathcal{R}(\rho_0)
    &= \rho_0 - IS(\rho_0) + R(-2\pi,\rho_0,I)
    &< \rho_0 - IS(\epsilon) + I^2r^\prime
    .
\end{align*}
Choosing \(I_5 \in ]0,I_4]\) so that 
\((I_5)^2r^\prime - I_5 S(\epsilon) \leq 0\) makes
\(P_\mathcal{R}(\rho_0) < \rho_0\) for all 
\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}[\).

Finally, choosing \(I^\star \in ]0,\min{I_3,I_5}]\) yields that
\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all \(\rho_0 \in D\), as
desired.
Looking at ~\eqref{eqn:rhohat-discrete}, this means all solutions of the
discrete-time system will flow away from the origin until it reaches a magnitude
\(\abs{\rho(n)} = \bar{\rho}\).
Hence, the constrained dynamics will eventually escape the set
\(\omega \subset \mathcal{R}\) defined by
\[
    \omega := \left\{(q_u,p_u) \in \Gamma
        \mid E(\pi,0) < E(q_u,p_u) < E(0,\bar{\rho} \right\}
    ,
\]
which is displayed in Figure \ref{fig:acrobot-little-omega}.

We conclude that there exists \(I > 0\) small enough that the constraint
~\eqref{eqn:acrobot-constraint} is injecting energy on \(\omega\).

\begin{figure}
    \centering
    \includestandalone[width=0.5\textwidth]{images/acrobot_little_omega}
    \caption{The set \(\omega \subset \mathcal{R}\) on which the acrobot will gain energy.
    }
    \label{fig:acrobot-little-omega}
\end{figure}

By the same arguments presented in this section, the Poincar\'{e} section
satisfies \(\abs{P_\mathcal{R}(\rho_0)} < \abs{\rho_0}\) when \(I < 0\).
In this case, the constraint dissipates energy on \(\omega\).

\subsection{Energy Gain on \(\Omega\)}
Define \(E_\pi\) to be the energy level set associated with \(E(pi,0)\).
We will prove the acrobot gains energy on the set 
\begin{align*}
    \Omega &= \left\{(q_u,p_u) \in \Gamma
    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,\\
    &= \mathcal{O} \cup E_\pi \cup \omega
    .
\end{align*}
There exists some control input \(I_1 > 0\) for which the acrobot is gaining
energy on \(\mathcal{O}\). 
Hence, there are no closed orbits in \(\mathcal{O}\) and almost all orbits flow
towards \(E_\pi\).
Likewise, there exists some \(I_2 > 0\)for which the acrobot is gaining energy
on \(\omega\), so there are no closed orbits in \(\omega\) and almost every
orbit is flowing away from \(E_\pi\).
Choosing \(I^\star \in ]0,\min{I_1,I_2}]\) means the acrobot will gain energy on
both \(\mathcal{O}\) as well as \(\omega\).

We can prove energy gain on \(\Omega\) by showing two properties of the system.
First, that the set \(\Pi^+\) of initial conditions converging to  \((\pi,0)\)
is of measure zero in \(\Gamma\);
second, that \(E_\pi\) is not an invariant set and therefore is not equal to 
\(\Pi^+\).
If both these conditions are true, then orbits flowing from \(\mathcal{O}\) into
\(E_\pi\) will generally swing past \(E_\pi\) into \(\omega\).

To show the first condition, we will simply prove that \((\pi,0)\) is an
unstable equilibrium of the constrained dynamics.
Taking the jacobian of ~\eqref{eqn:acrobot-constrained-dynamics} at \((\pi,0)\)
yields
\[
    J = \begin{bmatrix}
        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
        3mgl & mgl\bar{q}_aI
    \end{bmatrix}
    ,
\]
which has the characteristic polynomial
\[
    \det\left(\lambda \Id{2} - J\right)
    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g
    .
\]
By Descartes' rule of signs, this polynomial has one root with positive real
part. 
Hence, the equilibrium \((\pi,0)\) is unstable.

To show the second condition, we evaluate \(\dot{E}(q_u,p_u)\) at the point
\((0,\pm\sqrt{60m^2gl^3}) \in E_\pi\).
MATLAB returns the value
\[
    \dot{E}(0,\pm\sqrt{60m^2gl^3}) 
    = \mp\frac{g}{5l}
    \sqrt{60m^2gl^3} \sin\left(\bar{q}_a \arctan(I \sqrt{60m^2gl^3})\right)
    .
\]
This is zero only if \(\bar{q}_a \arctan(I\sqrt{60m^2gl^3}) = \frac{\pi}{2}\), a
situation which occurs only when \(\bar{q}_a \geq 1\) at a single value of
\(I\).
Avoiding this value of \(I\) (if it exists) makes \(E_\pi\) a non-invariant set
as desired.

We know that almost all orbits starting in \(\omega\) will escape \(\Omega\).
Non-invariance of \(E_\pi\) and instability of \((\pi,0)\) implies that 
almost all orbits initialized in \(\mathcal{O}\) will reach \(E_\pi\) and
eventually swing into \(\omega\), rather than stay in \(E_\pi\) and converge to
\((\pi,0)\). 
We conclude that almost all orbits will depart the energy level set associated with
\(E(0,\bar{\rho})\), thereby escaping \(\Omega\) all together.
Hence, the acrobot is gaining energy on \(\Omega\).

If \(I < 0\), orbits in \(\mathcal{O}\) will converge to the origin and orbits in
\(\omega\) will flow towards \(E_\pi\).
However, the upright equilibrium \((\pi,0)\) will still be unstable, and
\(E_\pi\) will remain a non-invariant set. 
This means that almost all orbits in \(\omega\) will flow past \(E_\pi\) into
\(\mathcal{O}\), so the constrained system will be gaining energy in
negative-time. Thus, the constraint is dissipating energy on \(\Omega\).
